Covariate analysis of MSSM reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-03-21 code/Rmd/MAYO_REPROCESSED_TMM.Rmd
fixed reprocessed metadata issue with MSSM cohorts

# Conditional quantile normalisation
tmp.count = COUNT
genesToAnalyze = unique(intersect(rownames(GENE.GC)[!is.na(GENE.GC$percentage_gc_content)], 
                                  rownames(GENE.LEN)[!is.na(GENE.LEN$gene.length)]))
tmp.count = tmp.count[rownames(tmp.count) %in% genesToAnalyze,]
CQN.GENE_EXPRESSION = cqn(tmp.count,
                          x = GENE.GC[rownames(tmp.count), 'percentage_gc_content'],
                          lengths = GENE.LEN[rownames(tmp.count), 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)

# Initial normalisation of gene expression
ind = rownames(CQN.GENE_EXPRESSION$y) %in% CONTROL.PROBES$ensembl_gene_id 
EXPR = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
EXPR = EXPR[ind,]
PMI = PMI[colnames(EXPR),,drop=F]

@@

multiplot(plotlist = p, cols = 3)
```
### CPM Normalisation
Preprocess counts matrix using CQN (Conditional Quantile Normlaisation accounting for both GC content and gene length)

@@

### Normalisation (with NULL)
Initial normalisation usign cqn and dispersion control with voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# Normalise using TMM
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION)

>>>>

# Conditional quantile normalisation
tmp.count = COUNT
genesToAnalyze = unique(intersect(rownames(GENE.GC)[!is.na(GENE.GC$percentage_gc_content)], 
                                  rownames(GENE.LEN)[!is.na(GENE.LEN$gene.length)]))
tmp.count = tmp.count[rownames(tmp.count) %in% genesToAnalyze,]
VOOM.GENE_EXPRESSION = voom(tmp.count)

# Initial normalisation of gene expression
ind = rownames(VOOM.GENE_EXPRESSION$E) %in% CONTROL.PROBES$ensembl_gene_id 
EXPR = VOOM.GENE_EXPRESSION$E
EXPR = EXPR[ind,]
PMI = PMI[colnames(EXPR),,drop=F]

@@

multiplot(plotlist = p, cols = 3)
```
### CPM Normalisation
Preprocess counts matrix using TMM

@@

### Normalisation (with NULL)
Initial normalisation usign TMM and dispersion control with voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# Normalise using TMM
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION)

### 2017-03-21 code/Rmd/MSSM_REPROCESSED.Rmd
fixed reprocessed metadata issue with MSSM cohorts

# Get technical metadata
METADATA_TECH_ID <- 'syn6100548'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECH_ID
METADATA_TECH <- read.table(synGet(METADATA_TECH_ID)@filePath,sep=',',header=T, row.names=1) %>%
  group_by(sampleIdentifier) %>%
  top_n(1, -rRNA.rate) %>%
  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN) %>%
  unique

@@

METADATA <- METADATA %>%
  filter(!is.na(BrainRegion.Diagnosis), !is.na(PMI), !is.na(RIN))
```
Following samples were considered outliers and removed based on PCA and/or eucledian distance: 11311\_CER, 1923\_CER, 1923\_TCX, 11396\_TCX, 11294\_TCX, 11408\_TCX
```{r preprocess.data1}
# Match covariates to expression data
indToRetain = intersect(METADATA$sampleIdentifier, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$sampleIdentifier)

>>>>

# Get technical metadata
METADATA_TECH_ID <- 'syn6100548'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECH_ID
METADATA_TECH <- read.table(synGet(METADATA_TECH_ID)@filePath,sep=',',header=T, row.names=1) %>%
  group_by(sampleIdentifier) %>%
  top_n(1, -rRNA.rate) %>%
  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN) %>%
  unique %>%
  ddply(.(barcode), .fun = function(x){
    x = dplyr::mutate(x, batch = paste(batch, collapse = ''), RIN = sum(RIN, na.rm = T))  
  })

@@

METADATA <- METADATA %>%
  filter(!is.na(BrainRegion.Diagnosis), !is.na(PMI), !is.na(RIN))
```

```{r preprocess.data1}
# Match covariates to expression data
indToRetain = intersect(METADATA$sampleIdentifier, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$sampleIdentifier)
